Technology-based aids for people affected by Autism Spectrum Disorder (ASD)

### 2017-12-07 paper.tex
Added conclusions

With respect to these aims, this work is divided into two main sections: In \autoref{sec:diagnosis} we present some innovative tools and procedures to help diagnosing the syndrome in the first one or two years of a child; in \autoref{sec:thera_tech} we describe three \cri{vedere se poi il numero va bene} devices and methods useful for learning and improving cognitive skills.

@@

Moreover, although normally it is not considered as a diagnostic criteria, motor impairment associated with ASD may have a significant impact on daily life tasks and, since up to 80\% of children with autism show motor anomalies highly correlated with the severity of social deficiency, hand-eye coordination could be considered as a new early diagnostic strategy for ASD syndrome \cite{Casellato2017}. In fact, it has been suggested that sensory integration with bad timing brings to poor motor performance in children with ASD: the task of integrating multisensory input influences the forming of coherent perception, action coordination and planning. Nowadays, the normal procedure to assess this motor coordination deficiency is the clinical observation and self-report questionnaires, so a new non-invasive objective method to map the development of ASD such hand-eye motor performance, could be a powerful diagnostic tool.

@@

Figures \ref{fig:task1} and \ref{fig:task1_pT} show the results just described.

@@

Figure \ref{path} shows an example of trial with a TD child.

@@

The experiment results can be summarized with a few crucial conclusions:
\begin{itemize}
\item the first task showed how children with \gls{asd} took longer to learn the task execution correctly, even though once the process had been assimilated the execution was correct, this meaning that these children showed a delayed ability to learn from sensory prediction errors; moreover the ASD child had more variability in motor execution, this suggesting more difficulty in maintaining performance consistency, and the gaze pattern was sticking less to the hand pattern if compared to the \gls{td} children.
\item the second task showed how children with ASD obtained performance slightly worse than the TD children, but anyway they were able deploy the information regarding the previously assimilated path to improve the performance. This suggests that gls{asd} children are able to take advantage of the availability of the whole motor plan and supports the hypothesis that simple motor planning is intact but there is a diminished use of feedback.
\end{itemize}

@@

The signal processing passage consisted in low-pass filtering at 100 Hz, subsequent sampling at 256 Hz and digitization through 12-bit ADC. Finally. EEG signals were divided into 3 artifact-free epochs, subsequently filtered trough a Hamming window FIR band-pass filter.

@@

Looking at the whole system it can be stated that this tool can allow a very useful didactic training procedure for people with \gls{asd}, and its potential affordability for normal families and relatively easy use make it a powerful solution in the research field  of \gls{asd}. \cri{forse devo essere più specifica nella parte in cui descrivo come il sistema as adatta le task all'ansia?}

\section{Conclusions}\label{sec:conclusions}

>>>>

With respect to these aims, this work is divided into two main sections: In \autoref{sec:diagnosis} we present some innovative tools and procedures to help diagnosing the syndrome in the first one or two years of a child; in \autoref{sec:thera_tech} we describe two devices and methods useful for learning and improving cognitive skills.

@@

Moreover, although normally it is not considered as a diagnostic criteria, motor impairment associated with ASD may have a significant impact on daily life tasks and, since up to 80\% of children with autism show motor anomalies highly correlated with the severity of social deficiency, hand-eye coordination could be considered as a new early diagnostic strategy for ASD syndrome \DP{Ho tolto la reference al primo paper, mi sembrava ridondante}. In fact, it has been suggested that sensory integration with bad timing brings to poor motor performance in children with ASD: the task of integrating multisensory input influences the forming of coherent perception, action coordination and planning. Nowadays, the normal procedure to assess this motor coordination deficiency is the clinical observation and self-report questionnaires, so a new non-invasive objective method to map the development of ASD such hand-eye motor performance, could be a powerful diagnostic tool.

@@

\autoref{fig:task1} and \autoref{fig:task1_pT} show the results just described.

@@

Figure \autoref{path} shows an example of trial with a TD child.

@@

The experiment results can be summarized with a few crucial conclusions:
\begin{itemize}
\item the first task showed how children with \gls{asd} took longer to learn the task execution correctly, even though once the process had been assimilated the execution was correct, this meaning that these children showed a delayed ability to learn from sensory prediction errors; moreover the ASD child had more variability in motor execution, this suggesting more difficulty in maintaining performance consistency, and the gaze pattern was sticking less to the hand pattern if compared to the \gls{td} children.
\item the second task showed how children with ASD obtained performance slightly worse than the TD children, but anyway they were able to deploy the information regarding the previously assimilated path to improve the performance. This suggests that \gls{asd} children are able to take advantage of the availability of the whole motor plan and supports the hypothesis that simple motor planning is intact but there is a diminished use of feedback.
\end{itemize}

@@

The signal processing passage consisted in low-pass filtering at 100 Hz, subsequent sampling at 256 Hz and digitization through 12-bit ADC. Finally, EEG signals were divided into 3 artifact-free epochs, subsequently filtered trough a Hamming window FIR band-pass filter.

@@

Looking at the whole system it can be stated that this tool can allow a very useful didactic training procedure for people with \gls{asd}, and its potential affordability for normal families and relatively easy use make it a powerful solution in the research field  of \gls{asd}. \cri{forse devo essere più specifica nella parte in cui descrivo come il sistema as adatta le task all'ansia?}
\DP{No secondo me quando hai spiegato che il sistema si adatta all'ansia e come rileva l'ansia, siamo apposto! Come hai fatto tu va benissimo!}

\section{Conclusions}\label{sec:conclusions}
All the presented experiments had reach good results while applied to subjects with \gls{asd}. This proofs that new technologies can be used, together with a specialized therapist, in the medical fields to prognosticate or manage the \gls{asd}.

Unfortunately, before these tools are normally used in clinics and hospitals, it must takes a non-negligible amount of time and a lot of tests, but in a \textit{not-so-far} future, these techniques can help us to improve the life of the people affected by \gls{asd} and maybe they will allow us a better understanding of this desease.
